---
title: "News"
author: "Troy Hall, Sam Galloway, Rees Braam, Sidh Kulgod"
date: "Aproil 8, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, warning=FALSE}
set.seed(1337)
library(tidyverse)
library(dplyr)
library(readr)
```

```{r, cache=TRUE, message=FALSE, warning=FALSE}
# Read the datasets we specified

# Online News Popularity from https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity
OnlineNewsPopularity <- rbind(read_csv("C:/Users/smg635/Downloads/data/OnlineNewsPopularityTraining.csv"), read_csv("C:/Users/smg635/Downloads/data/OnlineNewsPopularityTest.csv"))

# All the news from https://www.kaggle.com/snapcrack/all-the-news
AllTheNews <- read_csv("C:/Users/smg635/Downloads/data/all-the-news.csv")

# All the news from https://www.kaggle.com/uciml/news-aggregator-dataset
NewsAggregatorDataset <- read_csv("C:/Users/smg635/Downloads/data/news-aggregator-dataset.csv")
```

```{r}
head(OnlineNewsPopularity)

head(AllTheNews)

head(NewsAggregatorDataset)
```


```{r}
# AllTheNews <- AllTheNews %>% select(1:month, day, day_of_week, url:polarity) # Switches order of columns
AllTheNews <- rename(AllTheNews, text = content)
colnames(AllTheNews)
```

```{r}
colnames(NewsAggregatorDataset)
```

### Filter the rows in *OnlineNewsPopularity* where there was an error getting the text and rename the subjectivity and polarity columns
```{r}
# Remove the 22 NA rows from OnlineNewsPopularity

sum(is.na(OnlineNewsPopularity$text) == TRUE)
OnlineNewsPopularity <- filter(OnlineNewsPopularity, is.na(OnlineNewsPopularity$text) == FALSE)
sum(is.na(OnlineNewsPopularity$text) == TRUE)

OnlineNewsPopularity <- rename(OnlineNewsPopularity, subjectivity = global_subjectivity, polarity = global_sentiment_polarity)
```

### Filter out the rows where date had an issue being parsed, maybe fixable at a later date
```{r}
nrow(filter(AllTheNews, AllTheNews$day_of_week == -1))
AllTheNews <- filter(AllTheNews, AllTheNews$day_of_week != -1)
nrow(filter(AllTheNews, AllTheNews$day_of_week == -1))
```

### Add columns for year, month, and day and format date columns as date
```{r}
AllTheNews$date <- as.Date(AllTheNews$date)
NewsAggregatorDataset$date <- as.Date(NewsAggregatorDataset$date)
OnlineNewsPopularity$date <- as.Date(OnlineNewsPopularity$date)

NewsAggregatorDataset$day <- as.numeric(format(NewsAggregatorDataset$date, format = "%d"))
NewsAggregatorDataset$month <- as.numeric(format(NewsAggregatorDataset$date, format = "%m"))
NewsAggregatorDataset$year <- as.numeric(format(NewsAggregatorDataset$date, format = "%Y"))

OnlineNewsPopularity$day <- as.numeric(format(OnlineNewsPopularity$date, format = "%d"))
OnlineNewsPopularity$month <- as.numeric(format(OnlineNewsPopularity$date, format = "%m"))
OnlineNewsPopularity$year <- as.numeric(format(OnlineNewsPopularity$date, format = "%Y"))
```

```{r}
colnames(AllTheNews)
```

```{r}
colnames(OnlineNewsPopularity)
```

```{r}
OnlineNewsPopularity$day_of_week <- 0 * OnlineNewsPopularity$weekday_is_monday + 1 * OnlineNewsPopularity$weekday_is_tuesday + 2 * OnlineNewsPopularity$weekday_is_wednesday + 3 * OnlineNewsPopularity$weekday_is_thursday + 4 * OnlineNewsPopularity$weekday_is_friday + 5 * OnlineNewsPopularity$weekday_is_saturday + 6 * OnlineNewsPopularity$weekday_is_sunday
```

```{r}
NewsAggregatorDataset <- rename(NewsAggregatorDataset, id = ID, title = TITLE, url = URL, publisher = PUBLISHER, category = CATEGORY, story = STORY, hostname = HOSTNAME, timestamp = TIMESTAMP)

NewsAggregatorDataset$text <- NA
NewsAggregatorDataset$subjectivity <- NA
NewsAggregatorDataset$polarity <- NA
```

```{r}
AllTheNews2 <- AllTheNews
```



```{r}
AllTheNews <- select(AllTheNews, title, date, year, month, day, day_of_week, url, text, subjectivity, polarity)
OnlineNewsPopularity <- select(OnlineNewsPopularity, title, date, year, month, day, day_of_week, url, text, subjectivity, polarity)
NewsAggregatorDataset <- select(NewsAggregatorDataset, title, date, year, month, day, day_of_week, url, text, subjectivity, polarity)
```

```{r}
news = rbind(AllTheNews, OnlineNewsPopularity, NewsAggregatorDataset)
#news = rbind(AllTheNews, OnlineNewsPopularity, NewsAggregatorDataset)
```

```{r}
news <- mutate(news, numwords = map_int(str_split(text,boundary("word")),length),
               lenwords = map(str_split(text,boundary("word")), ~ str_count(., "[A-z]")))
```
###Plot the length and number of words

###Sets date variables as factors
```{r}
news$month <- as.factor(news$month)
news$numwords <- as.numeric(news$numwords)
news$day_of_week <- as.factor(news$day_of_week)
news$year <- as.factor(news$year)
```

###Makes a version of news we can look at for number of words, as it excludes those with too few words
```{r}
news2 <- filter(news,(news$numwords >= 5) & (news$numwords <= 5000))
```

###Plotting the number of words by weekday, year, and month

```{r}
ggplot(news2,aes(x = year, y = numwords)) + geom_boxplot(fill = "lightsteelblue1")  + theme_bw()
```
